---
title: "Bluemussel"
output: html_document
---

## 1. Packages  
```{r}

library(tidyverse)
library(readxl)
library(glue)

# Load function `get_lengde_from_file`
source("01_Bluemussel_lengths_excel_functions.R") 

```

## 2. Data   


### Data folders  
```{r}

folders_2014_2019 <- c(
  "K:/Prosjekter/Sjøvann/JAMP/2014/opparbeiding/blaaskjell",
  "K:/Prosjekter/Sjøvann/JAMP/2015/Opparbeiding_biota/Blaaskjell",
  "K:/Prosjekter/Sjøvann/JAMP/2016/opparbeiding biota/blåskjell",
  "K:/Prosjekter/Sjøvann/JAMP/2017/opparbeiding biota/blaaskjell",
  "K:/Prosjekter/Sjøvann/JAMP/2018/opparbeiding biota/Blåskjell" ,
  "K:/Prosjekter/Sjøvann/JAMP/2019/opparbeiding biota/Blåskjell"
)

```

### Station names and coordinates  
```{r}

data_stations <- readxl::read_excel("../Milkys2_pc/Files_to_Jupyterhub_2019/Kartbase_edit.xlsx")

```


## 3. Test read one file  

### a. Files
```{r}

year <- 2018
datafolder <- folders_2014_2019[year - 2013]
fn <- dir(datafolder)
fn[1]
fn <- fn[!grepl("~",fn)]  


```
### b. Read sample no 1
```{r}

fn_full <- glue("{datafolder}/{fn[1]}")
# fn_full
dat <- read_excel(fn_full) %>% as.data.frame()
dat[18:27,1:4]
lengder <- bind_rows(data.frame (mm=20:29, antall=as.numeric(dat[18:27,2])),
                   data.frame (mm=30:39, antall=as.numeric(dat[18:27,3])),
                   data.frame (mm=40:49, antall=as.numeric(dat[18:27,4])))
lengder <- lengder %>%
  mutate(antall = case_when(
    is.na(antall) ~ 0,
    !is.na(antall) ~ antall
  ))

```


## 4. Test function `get_lengde_from_file`   

### a. 10A2_skallneset.xlsx in 2014 - a file with pretty unusual structure  
```{r}

# 10A2_skallneset.xlsx in 2014 - a file with pretty unusual structure 
#  (headings for each sample)
year <- 2014
datafolder <- folders_2014_2019[year - 2013]
fn <- dir(datafolder)
# fn[1]
fn <- fn[!grepl("~",fn)]  
fn[1]

# debugonce(get_lengde_from_file)

# Test  
df <- get_lengde_from_file(
  folder = datafolder,
  filename = fn[1],
  stationcode = "10A2",
  year = year
  )

ggplot(df , aes(x = Length, y = Number, fill = as.factor(Sampleno))) +
  geom_col()


```


### b. I714_Brevik in 2018 - 4 length groups per sample   
```{r}

# I714_Brevik in 2018 - 4 length groups per sample
year <- 2018
datafolder <- folders_2014_2019[year - 2013]
fn <- dir(datafolder)

# debugonce(get_lengde_from_file)

# Test  
df <- get_lengde_from_file(
  folder = datafolder,
  filename = "I714_Brevik kirke_2018.xlsx" ,
  stationcode = "10A2",
  year = year
  )

ggplot(df , aes(x = Length, y = Number, fill = as.factor(Sampleno))) +
  geom_col() +
  facet_wrap(vars(Sampleno))


```



## 5. Make file for file names and station names  

### Make file to be edited by hand   
Forgot year so that was also added manually
```{r}

make_filename_data <- function(year, folders, firstyear = 2014){
  datafolder <- folders_2014_2019[year - firstyear + 1]
  fn <- dir(datafolder)
  fn <- fn[!grepl("~",fn)]    # opened files
  tibble(
    Folder = datafolder,
    Filename = fn,
    Station_code = stringr::str_extract(fn, "[^_]+")
  )
}

df_files <- 2014:2019 %>% 
  purrr::map_dfr(
    make_filename_data, folders = folders_2014_2019, firstyear = 2014
  )

writexl::write_xlsx(df_files, "Data/01_df_files_to_be_edited.xlsx")

```

### Read edited file  
```{r}

df_files_all <- readxl::read_excel("Data/01_df_files_edited.xlsx")

df_files <- df_files_all %>%
  filter(is.na(Not_excel))

```


## 6. Get all data   

### Test with 2 files  
```{r}

test <- 1:2 %>%
  map_dfr(
    ~ suppressWarnings(
      get_lengde_from_file(
        folder = df_files$Folder[.],
        filename = df_files$Filename[.],
        stationcode = df_files$Station_code[.],
        year = df_files$Year[.],
        trace = TRUE
      )
    )
  )

ggplot(test , aes(x = Length, y = Number, fill = as.factor(Sampleno))) +
  geom_col() +
  facet_wrap(vars(Station))

```

### Get all data as list  
```{r}

# Make 'safe' version
# The safe version will always return a list of two elements,
#   'result' = the data frame from one file, if everything worked
#   'error' = NULL if everything worked
get_lengde_from_file_safe <- safely(get_lengde_from_file)

dat_list <- 1:nrow(df_files) %>%
  map(
    ~ suppressWarnings(
      get_lengde_from_file_safe(
        folder = df_files$Folder[.],
        filename = df_files$Filename[.],
        stationcode = df_files$Station_code[.],
        year = df_files$Year[.],
        trace = FALSE
      )
    )
  )


# Transpose the list, so we get a list of two elements:
#   'result' = a list of data frames (for those files that worked)
#   'error' = a list of error messages (NULL for those files that worked)
dat_list <- purrr::transpose(dat_list)

```

### Pick those that worked   
And check those that didn't    
```{r}

is_ok <- dat_list$error %>% map_lgl(is.null)
sum(!is_ok)

dat_list$error[!is_ok]

files_with_errors <- 
  df_files[!is_ok,]


```
### Put data together  
```{r}

dat <- dat_list$result[is_ok] %>% bind_rows()

```


### Add station coordinates and names  
```{r}

n1 <- nrow(dat)
dat <- dat %>%
  left_join(data_stations, by = c("Station" = "STATION_CODE"))

n2 <- nrow(dat)

if (n2 != n1){
  stop("Number of rows changed - stations in 'data_stations' are not unique?")
}  


```


## 7. Save raw data  
```{r}

# Reorder columns
dat <- dat %>%
  select(Station, Station_name, Lat, Long, Year, Sampleno, Length, Number)

# Save
saveRDS(dat, "Data/01_Bluemussel_lengths_excel.rds")
writexl::write_xlsx(dat, "Data/Bluemussel lengths 2014-2019 raw.xlsx")

if (FALSE){
  # Read back from file if necessary
  dat <- readRDS("Data/01_Bluemussel_lengths_excel.rds")
}

```

## 8. Save summarised data  


### Summarised on sample level  
3 lines per station per year  
```{r}

dat_summ1 <- dat %>%
  filter(!is.na(Number)) %>%
  group_by(Station, Year, Sampleno, Length) %>%
  summarise(Number = sum(Number)) %>%
  group_by(Station, Year, Sampleno) %>%
  summarise(
    Mean_length = weighted.mean(Length, Number),
    a = sum(Number*((Length - Mean_length)^2)),
    b = (n()-1)/n()*sum(Number),
    SD_length = sqrt(a/b),
    Min_length = min(Length),
    Max_length = max(Length),
    Sample_size = sum(Number),
    .groups = "drop") %>%
  select(-a, -b)

# Save
writexl::write_xlsx(dat_summ1, "Data/Bluemussel lengths 2014-2019 summarised sample.xlsx")

```
### Summarised on sample level  
1 lines per station per year  
```{r}

dat_summ2 <- dat %>%
  filter(!is.na(Number)) %>%
  group_by(Station, Year, Length) %>%
  summarise(Number = sum(Number)) %>%
  group_by(Station, Year) %>%
  summarise(
    Mean_length = weighted.mean(Length, Number),
    a = sum(Number*((Length - Mean_length)^2)),
    b = (n()-1)/n()*sum(Number),
    SD_length = sqrt(a/b),
    Min_length = min(Length),
    Max_length = max(Length),
    Sample_size = sum(Number),
    .groups = "drop") %>%
  select(-a, -b)

# Save
writexl::write_xlsx(dat_summ2, "Data/Bluemussel lengths 2014-2019 summarised station.xlsx")

```


## 9. Test plots    


### Means  
```{r}

ggplot(dat_summ2, aes(Year, Mean_length)) +
  geom_point() +
  geom_line() +
  facet_wrap(vars(Station))

```

```{r}

ggplot(dat_summ, aes(Year, Mean_length, color = Station)) +
  geom_point() +
  geom_line()

```
### SD
```{r}

gg <- ggplot(dat_summ, aes(Year, SD_length, color = Station)) +
  geom_point() +
  geom_line()

plotly::ggplotly(gg)

```

### Single stations  

```{r}

plot_station <- function(station){
  dat %>%
    filter(Station %in% station & !is.na(Number)) %>%
    group_by(Station, Year, Length) %>%
    summarise(Number = sum(Number), .groups = "drop") %>%
    ggplot(aes(x = Length, y = Number)) +
    geom_col(width = 1) +
    facet_wrap(vars(Year)) +
    labs(title = station)
}

plot_station("I131")
plot_station("76A2")
plot_station("10A2")

```






