---
title: "02_Bluemussel_Access"
author: "DHJ"
date: "10 11 2020"
output: html_document
---

Reading data stored in Access. For details, see README.   
Saved data are copied to `K:/Avdeling/214-Oseanografi/DHJ/Data/Contaminants/Marine/Milkys`  

## 1. Packages  
```{r}

library(tidyverse)
library(readxl)
library(glue)

```


## 2. Data   

### Data folder
```{r}

datafolder <- "K:/Avdeling/214-Oseanografi/DHJ/Data/Contaminants/Marine/Milkys"

```

### Station names and coordinates  
```{r}

data_stations <- readxl::read_excel(glue("{datafolder}/Kartbase_edit.xlsx"))

```

JAMP_Rapportering.mdb downloaded from `K:/Avdeling/Mar/NOG/JMGdata/Databaser` 27.08.2015   

### Read files  
```{r}

# Station/year level data (one line per year * station * species)
df_station <- read_csv2(
  glue("{datafolder}/Access/dbo_fish_mas.txt"), 
  col_types = cols(
  .default = col_character(),
  myear = col_double(),
  seqno = col_double(),
  cntry = col_double(),
  cruis = col_double(),
  sdate = col_double(),
  latitude = col_double(),
  longitude = col_double(),
  latdg = col_double(),
  latmi = col_double(),
  latmf = col_double(),
  londg = col_double(),
  lonmi = col_double(),
  lonmf = col_double(),
  noins = col_double(),
  wadep = col_double(),
  astsa = col_double(),
  ptsrc = col_double(),
  BEM = col_double()
))

# Sample level data  
df_sample <- read_csv2(
  glue("{datafolder}/Access/dbo_fish_spe.txt"), 
  col_types = cols(
  .default = col_double(),
  rlabo = col_character(),
  inorb = col_character(),
  sexco = col_character(),
  cones = col_character(),
  agdet = col_character(),
  seorb = col_character(),
  cmnt1 = col_character(),
  cmnt2 = col_character(),
  cmnt3 = col_character(),
  cmnt4 = col_character(),
  cmnt5 = col_character())
)

```
### Check duplicates  
```{r}

dupl <- df_station %>%
  group_by(rlabo, myear, seqno) %>%
  mutate(n = n()) %>%
  filter(n > 1) %>%
  arrange(myear, seqno)

# View(dupl)


```

### Check data sets 2010 and 2014     
```{r}

df_station %>%
  filter(rlabo == "NIVA" & myear == 2010) %>%
  xtabs(~dataset_id, .)

df_station %>%
  filter(rlabo == "NIVA" & myear == 2014 & jmpst == "35A") %>%
  xtabs(~dataset_id + cmnt1, .)

```

### Filter df_station   
To avoid duplication  
```{r}

sel <- with(df_station, rlabo == "NIVA" & myear == 2010 & dataset_id %in% c("O 10267","O 10269","O 10271"))
sum(sel)
df_station <- df_station[!sel,]

sel <- with(df_station, rlabo == "NIVA" & myear == 2014 & jmpst == "35A" & dataset_id == "LO 27487-2")
sum(sel)
df_station <- df_station[!sel,]

```

### Check duplicates again 
```{r}

dupl <- df_station %>%
  group_by(rlabo, myear, seqno) %>%
  mutate(n = n()) %>%
  filter(n > 1) %>%
  arrange(myear, seqno)

nrow(dupl)

```

## 3. Make 'df_bluemussel'  

### Join data
```{r}

n1 <- nrow(df_sample)
df_sample2 <- df_sample %>%
  left_join(df_station, by = c("rlabo", "myear", "seqno"))

n2 <- nrow(df_sample2) 
if (n2 != n1){
  stop("Number of rows changed - key columns in 'df_station' are not unique?")
}  

# Filter data 
df_bluemussel <- df_sample2 %>%
  filter(speci == "MYTI EDU" & myear > 1949) # 1949 is just mock data anyway

```


### Add station coordinates (fixed) and names  
```{r}

n1 <- nrow(df_bluemussel)
df_bluemussel <- df_bluemussel %>%
  left_join(data_stations, by = c("jmpst" = "STATION_CODE")) %>%
  rename(Lat_fixed = Lat,
         Long_fixed = Long)

n2 <- nrow(df_bluemussel)
if (n2 != n1){
  stop("Number of rows changed - stations in 'data_stations' are not unique?")
}  


```

## 4. Some statistics   

* Number (noinp) is lacking for ca 12%  
* Length is lacking for ca 15%, weight (wtmea) is lacking for 86% 
* Shell weight (shlwt and shlwtx) is lacking for 24-28%, tissue weight (tiswtx) for 81%    
* latitude, longitude lacking for 81%, but latdg, londg lacking for 34%  
* both latitude, longitude and latdg, londg lacking for 17%  

```{r}

# xtabs(~is.na(jmpst), df_bluemussel)
# apply(is.na(df_bluemussel), 2, mean)  

vars <- c("noinp", "lnmin", "lnmax", "lnmea", "lnstd", "wtmin", "wtmax", "wtmea", "wtstd", 
          "shlwt", "shlwtx", "tiswtx")
apply(is.na(df_bluemussel[vars]), 2, mean)  

xtabs(~myear, df_bluemussel %>% filter(!is.na(lnmea)))

vars <- c("latitude", "longitude", "latdg", "latmi", "latmf", "londg", "lonmi", "lonmf")
apply(is.na(df_bluemussel[vars]), 2, mean)  

sel <- with(df_bluemussel, is.na(latitude) & is.na(latdg))
mean(sel)

```
`
## Save + export  
```{r}

# names(df_bluemussel) %>% dput()
vars <- c("jmpst", "Station_name", "Lat_fixed", "Long_fixed",   # station codes (jmpst) and positions as 'defined' in 2019
          "myear", "sdate", "stime",                            # year and date
          "rlabo", "seqno",                                     # key columns in the original data
          "subno",                                              # Sample number (usually 1-3) - corresponds to 'Sampleno' 2014-2019
          "speci",                                              # species code (Mytilus edulis)
          "latitude", "longitude", "latdg", "latmi", "latmf", "londg", "lonmi", "lonmf",  # position as recorded in the data *)
          "inorb", "noinp",                                     # noimp = sample size (number of mussels in the sample)
          "lnmin", "lnmax", "lnmea", "lnstd",                   # Length per mussel (min, max, mean and SD)
          "wtmin", "wtmax", "wtmea", "wtstd",                   # Weight per mussel (lacking in most cases)
          "shlwtx", "tiswtx", "qeorw",                          # Total weight of shell and tissue (usually lacking)
          "cmnt1.x", "cmnt2.x", "cmnt3.x", "cmnt4.x", "cmnt5.x",   # Comments for sampling that year
          "cmnt1.y", "cmnt2.y", "cmnt3.y", "cmnt4.y", "cmnt5.y")   # Saample-specific comments

# *) position variables: 
#   latitude = latitude with decimals
#   latdg, latmi, latmf = latitude given as whole degrees, minutes (1/60 degree) and the two first decimals of minutes (1/100 minute)
#   Note: these positions may differ from one year to another (at least in principle)

# Selected  columns for excel file
# For som reason, 
# writexl::write_xlsx(df_bluemussel[,vars], "Data/Bluemussel lengths 1981-2013 summarised.xlsx")  

# Save as csv (European format)
readr::write_csv2(df_bluemussel[,vars], "Data/Bluemussel lengths 1981-2013 summarised.csv")  
# To save this to excel: 1) Open an empty excel file; 2) Data:From text/csv , 
# 3) In the preview of the file, change "File Origin" to  "65001: Unicode (UTF-8)", 4) Load
# 5) Copy all, open yet another excel file, and "paste as values"


# All columns for rds file
saveRDS(df_bluemussel, "Data/02_df_bluemussel.rds")

```

## 6. Test plots  
```{r}

df_bluemussel %>%
  filter(jmpst == "30A") %>%
  ggplot(aes(myear, lnmea, color = noinp)) +
  geom_point() +
  facet_wrap(vars(jmpst))

```

```{r}

df_bluemussel %>%
  filter(!is.na(lnmea)) %>%
  group_by(jmpst) %>%
  mutate(n_years = length(unique(myear))) %>%
  filter(n_years >= 5) %>%
  ggplot(aes(myear, lnmea, color = noinp)) +
  geom_point(size = 0.5) +
  facet_wrap(vars(jmpst))

```


